---
title: "Introduction to whattheflux"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to whattheflux}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `whattheflux` package provides functions to parse static-chamber
greenhouse gas measurement files generated by a variety of instruments; 
compute flux rates using multi-observation metadata; and generate diagnostic
metrics and plots. It's designed to be easy to integrate into scientific workflows.

Load some sample data:

```{r setup}
library(whattheflux)

# Data from a LI-7810
f <- system.file("extdata/TG10-01087.data", package = "whattheflux")
dat <- wtf_read_LI7810(f)
dat[1:6, 1:9] # full raw dataset has 500+ rows and 25 columns
```

The analyzer data is basically a stream of timestamps and associated
greenhouse gas concentration measurements. For this to be useful, we need to 
associate it with metadata about the measurements we made: when they started,
how long they lasted, plot/treatment/collar information, etc. 

```{r metadata}
# Accompanying metadata
md <- system.file("extdata/TG10-01087-metadata.csv", package = "whattheflux")
metadat <- read.csv(md)

# Convenience function to fill in missing dead band and observation length info
metadat <- wtf_fill_metadata(metadat)

print(metadat)
```

The `wtf_metadata_match` function matches up the data, using the `TIMESTAMP`
columns that `wtf_read_LI7810` helpfully created when it read the file.

```{r matching, fig.width=8}
dat$metadat_row <- wtf_metadata_match(dat$TIMESTAMP, 
                                      metadat$Date, metadat$Start_time,
                                      metadat$Dead_band, metadat$Obs_length)
dat$Plot <- metadat$Plot[dat$metadat_row]
metadat$metadat_row <- seq_len(nrow(metadat))

# ...and plot
library(ggplot2)
p <- ggplot(dat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point()
print(p)
```

Some of these are clearly not correct--the measurement time seems to be shorter
then 60 seconds for the C, D, and E plots:

```{r circles, fig.width=8, echo=FALSE, message=FALSE}
library(lubridate)
p + annotate("rect", ymin = 455, ymax = 476, 
             xmin = ymd_hms("2022-10-27 10:39:30", tz = "EST"), 
             xmax = ymd_hms("2022-10-27 10:43:30", tz = "EST"), 
             color = "red", fill = NA, linewidth = 1.5)
```

In real life we'd probably want to correct the faulty metadata at its source.
Here, we change the values programmatically and re-match:

```{r matching2, fig.width=8}
metadat$Obs_length[3:5] <- c(30, 45, 45)
dat$metadat_row <- wtf_metadata_match(dat$TIMESTAMP, 
                                      metadat$Date, metadat$Start_time,
                                      metadat$Dead_band, metadat$Obs_length)
dat$Plot <- metadat$Plot[dat$metadat_row]

p %+% dat
```

That looks better!

